Brief notes on erosion

### 2019-04-22 docs/proposals/ErosionWhitePaper.tex
working on DOE white paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, from massive land formations to
centimeter-scale features and patterns. Though less visible, these same
forces are working at the very smallest scales, slowly deteriorating the
individual constituents of porous media (e.g.~soil, sand, or clay) or
biological structures like plaque and biofilms. Motivated by such
examples, we are interested in studying fluid-mechanical erosion in
Stokes flow---the most relevant regime for groundwater and bio-fluid
applications.

\begin{figure}[htp]
\includegraphics[width=\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical shear. The flow is horizontal
(left to right) and the 6 snapshots are evenly spaced in time. Color
represents vorticity, which provides a convenient way to visualize local
shear rates. Erosion not only diminishes the size of the bodies but also
alters their shapes considerably. A few well-defined channels develop in
the space between bodies.}
\end{figure}

In our recent publication~\cite{qua-moo2018}, we described a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  incompressible Stokes equations are solved with a boundary integral
  equation (BIE).  By using a BIE, the complex geometries are naturally
  handled, high-order accurate discretizations are deployed, and fast
  summation methods result in optimal complexity.

  \item {\bf Interface Tracking}: Instead of tracking the $(x,y)$
  coordinates of the boundary of the eroding bodies interface, we track
  the \thL coordinates, where $\theta$ is a local tangent angle, and $L$
  is the total length.  The governing equations of the \thL coordinates
  offers several advantages in the ability to numerically stabilize the
  evolving interface

  \item {\bf Regularizations of the Interface}: Erosion results in
  corners of the bodies at stagnation points of the flow.  We 
  use a regularization term and a smoothing operation to control the
  smoothness of the moving interface.

@@

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, additional numerical tools are required to simulate more
realistic scenarios such as bodies that erode and transport, and
extending our work to three dimensions.  Second, we will investigate
physically motivated questions regarding the geometry and flow inside an
eroding porous medium

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
In its current formulation, our bodies are fixed in space.  However, in
reality, eroding bodies are also transported by the flow.  To allow the
bodies to move, there are two additional tools that we will develop.
Bodies get close.

\paragraph{Transport of Bodies:} The PI is experienced with simulating
moving interfaces in Stokes flow.  For example, he has developed
high-order adaptive time stepping methods for simulating deformable
bodies (vesicles) in a viscous fluid~\cite{qua-bir2015b, qua-bir2014,
qua-2016}.   More recently, the PI has graduated a Ph.D.~student that
developed time stepping methods for rigid bodies in a viscous
fluid~\cite{bys-sha-qua2018}.  This work introduces a new time stepping
method

A recent Ph.D.~student from his group

\paragraph{Improved Quadrature:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Physics-Based Understanding of Viscous Erosion}
\todo[inline]{Characterizing anisotropy}
\todo[inline]{Coarse-grained transport (anomalous diffusion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{refs}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{wrapfigure}[24]{r}{0.7\textwidth}
\centering
\includegraphics[width=0.68\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} \em Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical erosion. The flow is
horizontal (left to right) and the 6 snapshots are evenly spaced in
time.  The local erosion rate on the bodies is proportional to the
magnitude of the vorticity which we compute in the fluid bulk (color
scale).  Erosion not only diminishes the size of the bodies but also
alters their shapes considerably. A few well-defined channels develop in
the space between bodies.}
\end{wrapfigure}

Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, including massive land formations
(mountains undergoing wind erosion) and watercourses (rivers and streams
undergoing hydraulic erosion).  Though less visible, erosion is working
at the very smallest scales, slowly deteriorating the individual
constituents of porous media (e.g.~soil, sand, or clay) or biological
structures like plaque and biofilms. Motivated by such examples, we are
interested in studying fluid-mechanical erosion in Stokes flow---the
most relevant regime for groundwater and bio-fluid applications.

In our recent publication~\cite{qua-moo2018}, we describe a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  incompressible Stokes equations are solved with a boundary integral
  equation (BIE).  By using a BIE, the complex geometries are naturally
  handled, high-order accurate discretizations are deployed, and fast
  summation methods result in optimal complexity.

  \item {\bf Interface Tracking}: Instead of tracking the $(x,y)$
  coordinates of the of the eroding interfaces, we track the {\thL}
  coordinates, where $\theta$ is a local tangent angle, and $L$ is the
  total interface length.  The governing equations of the {\thL}
  coordinates offers several advantages in the ability to numerically
  stabilize the evolving interface.

  \item {\bf Regularizations of the Interface}: Erosion results in
  numerically challenging corners of the bodies at stagnation points of
  the flow.  We use a regularization term and a smoothing operation to
  control the smoothness of the moving interface.

@@

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, additional numerical tools are required to simulate more
realistic scenarios such as denser packings, transport, and
three-dimensional erosion.  Second, we will investigate physically
motivated questions regarding the geometry and flow inside an eroding
porous medium.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
\begin{wrapfigure}[22]{r}{0.4\textwidth}
\centering
\input{figs/vort1.tikz}
\caption{\label{fig:vort1} \em The vorticity a vertical slice cutting
through a single body.  There are solid walls at $y = \pm 1$, and an
eroding body at $y = (-0.5,0.5)$.  The trapezoid rule gives large errors
near the solid walls, while the Barycentric quadrature formula results
in a smooth vorticity field.}
\end{wrapfigure}

In its current formulation, our bodies are fixed in space.  However, in
reality, eroding bodies are also transported by the flow.  To allow the
bodies to move, there are two additional tools that we will develop.
The PI is experienced with simulating moving interfaces in Stokes flow.
For example, he has developed high-order adaptive time stepping methods
for simulating deformable bodies (vesicles) in a viscous
fluid~\cite{qua-bir2015b, qua-bir2014, qua-bir2016}. More recently, the
PI and his former Ph.D.~student developed time stepping methods for
rigid body transport in a viscous fluid~\cite{bys-sha-qua2018}.  This
work extends a contact algorithm, originally developed by Lu et
al.~\cite{lu-rah-zor2017}, to accuratly simulate the dynamics of
nearly-touching rigid bodies.  The contact algorithm resolves errors
introduced by time stepping, but quadrature error introduced by the
integral equation formulation must also be addressed.  A postdoc in the
PI's group is extending a Barycentric near-singular integration scheme
developed by Barnett et al.~\cite{bar-wu-vee2015} to develop quadrature
formula to compute the shear stress and vorticity.  In
Figure~\ref{fig:vort1}, the vorticity is computed along a vertical slice
that cuts through a single eroding body.  The trapezoid rule is accurate
far from the solid walls ($y = \pm 1$) and the boundary of the eroding
body ($y = \pm 0.5$), but introduces large errors near the walls and
body.  However, the Barycentric quadrature rule is uniformly accurate
throughout the entire fluid domain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Physics-Based Understanding of Viscous Erosion}
\todo[inline]{Characterizing anisotropy}
\todo[inline]{Coarse-grained transport (anomalous diffusion)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\footnotesize{\bibliography{refs}}
